---
title: "THL Open API R package"
author: "Tuomo Nieminen, Leo Lahti"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
  rmarkdown::md_vignette:
    toc: true    
vignette: >
  %\VignetteIndexEntry{THL Open API R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{Cairo}
  \usepackage[utf8]{inputenc}  
---

The **openthl R package** provides tools to work with THL Open API.


## Installation

Install the current devel version in R:

```{r install, eval=FALSE}
devtools::install_github("ropengov/openthl")
```

Test the installation by loading the library:

```{r test, message=FALSE, warning=FALSE, eval=TRUE}
library(openthl)
```

We also recommend setting the UTF-8 encoding:

```{r locale, eval=FALSE}
Sys.setlocale(locale="UTF-8") 
```


## Examples

```{r}
library(openthl)

x <- getSubject("epirapo") # S3 classes "api_response" "thl_subject" 
datasets <- getDatasets(x)
cube <- getCube(datasets, index = 1L) # a very messy object for now
```

### S3 classes

These are preliminary and very much subject to change

**api_data_url**

list with elements:

- api: url to the api
- path: defines what to get from the api
- format: "json"

has a print method

**api_response**

list with elements:

- response (parsed json response: a list or a data frame)
- api_data_url: api_data_url Object
- status: status code of the response




## Licensing and Citations

This work can be freely used, modified and distributed under the open license specified in the [DESCRIPTION file](https://github.com/rOpenGov/openthl/blob/master/DESCRIPTION).

Kindly cite the work as follows

```{r citation, message=FALSE, eval=TRUE}
citation("openthl")
```



## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```
